---
title: "HW8"
author: "Nathan Mehta"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## chi sq 
```{r}
table_exper<-matrix(c(165,95,152,52,85,52,67,35,85,37),ncol=2,nrow=5,byrow=TRUE)
table_exper
```

```{r}
colnames(table_exper)<-c("n of patients", "n of nausea")
rownames(table_exper)<-c("placebo","chlor","dimen","pento 100","pento 150")
```

```{r}
chisq.test(table_exper)
```
Here we are testing to see if the multinomial distribution is consisitent across all the drugs. The null hypothesis, H0: all the drugs and the placebo follow the same multinoimial distribution with the same probabilities (probability of nausea ) 

alt hyp: all the drugs have different MN distributions with different probabilities of nausea. 

with chi squared statistic 8.683 we fail to reject the null. (difference between observed nausea levels and expected for each drug not extreme enough to reject under null)


```{r}
pnorm(0,mean= -1, sd=2)
```
## permutation method for different samples
h0: sample means of method A and B are equal
h1: sample means are different 
```{r}
meth_a<-c(79.98,80.04,80.02,80.04,80.03,80.03,80.04,79.97,80.05,80.03,80.02,80.00,80.02)
meth_b<-c(80.02,79.94,79.98,79.97,79.97,80.03,79.95,79.7)
df_29<-c(meth_a,meth_b)
obs_diff<-mean(meth_a)-mean(meth_b)
chosen_partitions<-1000
diff__samples<-NULL
for(i in 1:chosen_partitions){
  meth_a__<-sample(df_29,length(meth_a),TRUE)
  meth_b__<-sample(df_29,length(meth_b),TRUE)
  diff__samples[i]=mean(meth_a__)-mean(meth_b__)
}
pval<-sum(abs(diff__samples)>=abs(obs_diff))/chosen_partitions
pval
hist(diff__samples)
```
with a p-value of .02, the permutation method concludes that the samples have different means. This p-value is more extreme than that of the t-test. 
b)the mann-whitney is similar in the way that both look to show whether the two groups are identical in distribution. 
## dot plots
```{r}
library(tidyverse)
```

```{r}
df__<-`11.40`
df__%>%ggplot(aes(x=present))+geom_dotplot()
```
```{r}
df__%>%ggplot(aes(x=absent))+geom_dotplot()
```
## t tests
#b
```{r}
sdx<-sd(df__$present)
sdy<-sd(df__$absent)
var__x<-sdx**2
var__y<-sdy**2
sqrt((var__x/10) + (var__y/10))
mean(df__$present)
mean(df__$absent)
```
#c
```{r}
t.test(df__$present,df__$absent,var.equal = F)
```
p-val= 0.001995

## non parametric
```{r}
difference<-df__$present-df__$absent
wilcoxtest <- wilcox.test(difference)
pval_diff_nonpara <- wilcoxtest$p.value
pval_diff_nonpara
```
the assumptions are that the distributions of the weight gains for treatment and control groups (present and absent) do not follow normal distributions.

H0: treatment doesn't work (present dist=absent dist)
h1: treatment cause change in mean weight gain
## more t testing
#a
```{r}
males_temp<-bodytemp%>%filter(gender==1)%>%select(temperature)
males_temp<-males_temp$temperature
fem_temp<-bodytemp%>%filter(gender==2)%>%select(temperature)
fem_temp<-fem_temp$temperature
qqnorm(males_temp)
qqnorm(fem_temp)
```
t-test can be used because the two samples fro males and females follow the normal distribution quantiles. can assume that the male and fem populations follow normal temp distributions

```{r}
xbar<-mean(males_temp)
ybar<-mean(fem_temp)
xbar-ybar
length(males_temp)
length(fem_temp)
x_var<-sd(males_temp)**2
y_var<-sd(fem_temp)**2
sqrt((x_var/65)+(y_var/65))
```
#b
```{r}
males_rate<-bodytemp%>%filter(gender==1)%>%select(rate)
males_rate<-males_rate$rate
fem_rate<-bodytemp%>%filter(gender==2)%>%select(rate)
fem_rate<-fem_rate$rate
qqnorm(males_rate)
qqnorm(fem_rate)
```
t-test can be used because the two samples fro males and females follow the normal distribution quantiles. can assume that the male and fem populations follow normal rate distributions.

```{r}
xbar_rate<-mean(males_rate)
xbar_rate
ybar_rate<-mean(fem_rate)
ybar_rate
xbar_rate-ybar_rate
length(males_rate)
length(fem_rate)
x_ratevar<-sd(males_rate)**2
y_ratevar<-sd(fem_rate)**2
sqrt((x_ratevar/65)+(y_ratevar/65))
```

```{r}
t.test(males_rate,fem_rate,var.equal = F)
```
#c
```{r}
difference<-males_temp-fem_temp
wilcoxtest <- wilcox.test(difference)
pval_diff_nonpara <- wilcoxtest$p.value
pval_diff_nonpara
```
H0:females and males have same distr. of temp.
h1: they have different dist.
reject null with p-value stated above
```{r}
difference<-males_rate-fem_rate
wilcoxtest <- wilcox.test(difference)
pval_diff_nonpara <- wilcoxtest$p.value
pval_diff_nonpara
```
H0:females and males have same distr. of rates.
h1: they have different dist. of rates

fail to reject the null with p-value of .6
