---
title: "Mult Reg"
author: "Nathan Mehta"
date: "10/19/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```


```{r}
pnt.in<-read.table("Downloads/pnt_pat.txt",T)
pnt_pat <- tibble(pnt.in) %>%  
   filter(wabaq != "NULL", 
        ethnicgp %in% c("AA","C"), 
        dx %in% c("Conduction", "Broca")) %>%  
   select(registry, wabaq, dx, ethnicgp, gender) %>%    
 unique() 
```

```{r}
class(pnt_pat$wabaq)
pnt_pat
```

##2
```{r}
lm.aphasia<-lm(wabaq~dx*ethnicgp,data=pnt_pat)
summary(lm.aphasia)
```
##3
The predicted aphasia quotient of an African American individual with Broca's aphasia equals the intercept.  

The predicted aphasia quotient of an African American individual with Conduction aphasia is the model intercept + 18.574

The predicted aphasia quotient of a Caucasian with Broca's aphasia is the intercept + .488

The predicted aphasia quotient of a Caucasian with Conduction aphasia is the intercept + 18.574 + .488 -3.159
##4
```{r}
pnt_pat %>% filter(ethnicgp %in% c("AA","C"), 
                   dx %in% c("Conduction", "Broca")) %>%  
  ggplot() + 
  aes(x = dx, color = ethnicgp, group = ethnicgp, y = wabaq) + 
  geom_smooth(method = 'lm')
```
There is no significance in the interaction of ethnicgp and dx because of the colinearity showed by the graph above. The AA and C regression lines (regressing wabaq onto dx) show similar upward trends at a similar slope, which suggests that ethnicgp and aphasia type do not affect each other significantly.

##5
```{r}
pnt_pat <- tibble(pnt.in) %>%  
   filter(wabaq != "NULL", 
        gender %in% c("F","M"), 
            dx %in% c("Conduction", "Broca")) %>%  
   select(registry, wabaq, dx, ethnicgp, gender) %>%    
 unique() 
```

```{r}
lm.aphasia2<-lm(wabaq~dx*gender,data=pnt_pat)
summary(lm.aphasia2)
```

This model suggests an insignificant interaction between gender and aphasia type as predictors of aphasia severity, with a p-value of .669. Individuals classified as M experience on average a 1.6 decrease in severity when having Broca's type. Individuals classified as having Conduction type severity experience a 18 point increase in severity when they are female.