---
title: "T and rank sum"
author: "Nathan Mehta"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data
```{r}
y1.<-c(0.28,0.51,1.00,0.39,0.29,0.36,0.32,0.69,0.17,0.33)
qqnorm(y1.)
y2.<-c(0.30,0.39,0.63,0.68,0.38,0.21,0.88,0.39,0.51,0.32)
qqnorm(y2.)
y3.<-c(1.07,1.35,0.69,0.28,1.24,1.53,0.49,0.56,1.02,0.30)
qqnorm(y3.)
#all 3 plots follow a straight line closely, making normal approximations relevant for each group
```
```{r}
groups<-c(rep("T1",10),rep("T2",10),rep("T3",10))
dataset<-c(y1.,y2.,y3.)
df_tidy<-data.frame(groups,dataset)
oneway.test(dataset~groups,var.equal = TRUE)
```

## calculate SSB
```{r}
y1.mean<-mean(c(0.28,0.51,1.00,0.39,0.29,0.36,0.32,0.69,0.17,0.33))
```
```{r}
y2.mean<-mean(c(0.30,0.39,0.63,0.68,0.38,0.21,0.88,0.39,0.51,0.32))
```
```{r}
y3.mean<-mean(c(1.07,1.35,0.69,0.28,1.24,1.53,0.49,0.56,1.02,0.30))
```
```{r}
mean(y1.mean,y2.mean,y3.mean)
10*(.469-.434)+(10*(.853-.434))
```
```{r}
sa<-sd(c(0.28,0.51,1.00,0.39,0.29,0.36,0.32,0.69,0.17,0.33))
sb<-sd(c(0.30,0.39,0.63,0.68,0.38,0.21,0.88,0.39,0.51,0.32))
sc<-sd(c(1.07,1.35,0.69,0.28,1.24,1.53,0.49,0.56,1.02,0.30))
sa
sb
sc
```
```{r}
##SSW
(9*(sa**2) + 9*(sb**2) + 9*(sc**2))
```
```{r}
1-pf(22.7,2,11.72)
```
## nonparam:
```{r}
groups_<-c(rep(1,10),rep(2,10),rep(3,10))
total_data<-c(y1.,y2.,y3.)
kruskal.test(total_data~groups_)

```





## calculations pairwise
```{r}
sd(aj$V1)
sd(f2$V1)
sd(c57$V1)
ajmean<-mean(aj$V1)
f2mean<-mean(f2$V1)
c57mean<-mean(c57$V1)
```
```{r}
ajmean-f2mean
ajmean-c57mean
c57mean-f2mean
```
```{r}
2*(1-pt(5.493,df=131))
2*(1-pt(15.435,df=131))
2*(1-pt(7.3416,df=131))
```
## pair t tests
```{r}
type1<-c(1.7, 1.9, 6.1,12.5, 16.5, 25.1, 30.5, 42.1, 82.5)
type2<-c(13.6, 19.8, 25.2, 46.2, 46.2, 61.1)
type3<-c(13.4, 20.9, 25.1, 29.7, 46.9)
groups<-c(rep("T1",9),rep("T2",6),rep("T3",5))
typedata<-c(type1,type2,type3)
df_tidy<-data.frame(groups,typedata)
```
```{r}
pairwise.t.test(df_tidy$typedata,df_tidy$groups,p.adjust="none")
```
H0: the two means in a t-test are the same
H1: the two means are different
comparing .34,.81, and .53 with alpha/3=.017, since all the p-values are greater, this says that we fail to reject the null for all 3 t-tests. the means of each group is the same.
## nonparam:
```{r}
library(tidyverse)
```
```{r}
df_tidy<-df_tidy%>%mutate(ranks=rank(typedata))
mean_rank_df<-df_tidy%>%group_by(groups)%>% summarize(mean_rank=mean(ranks))
joined_df<-df_tidy%>%
  left_join(mean_rank_df)%>%
  mutate(tot_mean=mean(mean_rank))%>%
  mutate(diff_sq=(mean_rank-tot_mean)^2)
joined_df
```
```{r}
SSB<-sum(joined_df$diff_sq)
K<-(12/(20*(20+1)))*SSB
1-pchisq(K,df=2)
```
since p-val > .05, we accept the null that all three types have the same distribution

